### 2022-06-26 paper/Chapters/Chapter01.tex
fix bugs in training evaluation
update ag model

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{From 3D to 2D}

\section{Kinect and Surface Normal}

As a popular and affordable new type of depth sensor, Kinect, had attracted a great focus to computer vision research in the last decades. It simultaneously captured the grayscale or RGB image and depth map with a high resolution. The depth map can further convert to structured point clouds with known intrinsic calibration parameters, which can use in many applications.

%% why surface normal is important
Surface Normal is one of the most worthy pieces of information that can infer from point clouds, which applies in many practical applications, such as augmented reality and robotics. Some tasks like shading and surface reconstruction require normal as one of the inputs. However, due to the lack of accuracy of the sensors, the  surface normal inference from depth maps/point clouds has many challenges.

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\section{Standard Methods}

%%%%%%%%%%%%%%%%%%% talk about the reason of this thesis %%%%%%%%%%%%%%%%%%

%% standard method for normal inference is insufficient, input is sparse, not robust enough
Standard methods compute normals from the point cloud using neighboring information in image space or from a single grayscale image using use Shape from Shading \cite{SFS}. The first method assumes that the neighbors of the points locate on the same plane. This method performs well with a well-chosen window size. However, the drawbacks are that the algorithm is highly noise sensitive. It is weak in handling missing pixels, which is a common issue in the input data. The second method depends on the correct information about the light source. Errors may occur in regions with inter-reflections in the 3D measurement.

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\section{Deep Learning based Method}
%% CNN based methods
Recently, deep learning based method \cite{yolov3}, \cite{efficientDet} achieved a great succeed for image processing. 
These network architectures use a batch of RGB/Grayscale images as input and usually employ for classification problems. Usually, the images are convoluted with a convolutional layer and downsampling with pooling layers. The outputs of the network consist of a single value to represent the index of the corresponding class \cite{efficientDet}, or with a set of values to represent the position of bounding boxes.\cite{yolov3}.

However, in many other vision tasks, like normal map inference, the output is demanded as the same shape as the input. Instead of predicting one or several classes for the whole input matrix, the class for each pixel requires for prediction. In this case, the traditional network architecture is not suitable anymore.

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\section{Challenges of Normal Inference}

%% talk about upsampling 
Recently, deep learning-based methods are used in computer vision tasks such as image segmentation, image inpainting, and depth density enhancement. In these methods, multiple architectures have proposed with an upsampling section. In this case, the output can be designed to have the same shape as the input or slightly smaller. Nevertheless, the resolution is similar to the input image. The normal Inference task has a similar pipeline compared to these tasks.

%% there exists space for improvement 
On the other hand, the point cloud data provided by Kinect or similar RGB-D, and LiDAR sensors are only semi-dense. A huge amount of missing pixels distribute all over the images, and some of the regions leave complete empty holes. This situation imposes another challenge on normal inference.

The depth image is incomplete, however, the depth sensor is able to capture grayscale texture images, which are typically fully dense due to their passive nature. Furthermore, if the texture image is already illuminated by strong directional light of a video projector, whose position is known, then there should exist theoretical relations between light direction, normal direction, and grayscale image. Then the normal can be inferred better using the given image information and depth map.  Based on the grayscale image and corresponding semi-dense depth maps, a CNN model can be designed to infer the normal map, which can give more density and robust results comparing to the standard algorithm.

%% add noise image
\begin{figure}[th]
	\centering
	\includegraphics[width=.4\textwidth]{Figures/kinect-depth.png}
	\decoRule
	\caption{A captured depth map via infrared sensors. Pixels that far away represent by light colors, otherwise by dark colors. The black dots are the depths that failed to be detected.}
\label{fig:depth-map-with-noise}
\end{figure}

%% add noise image

\begin{figure}[th]
	\centering
	\includegraphics[width=.4\textwidth]{Figures/real-normal.png}
	\decoRule
	\caption{Semi-dense Normal Map calculated from depth map using a standard method}
\label{fig:standard-normal-inference}
\end{figure}

%%%%%%%%%%%%%%%%%% talk about the main work in this thesis  %%%%%%%%%%%%%%%%%%
\section{Main Works of this thesis}
In this thesis, we found a solution for the problems mentioned above and proposed a novel deep learning architecture for surface normal inference. A network named Albedo Gated Normal Inference Network(AlGaN) is proposed to infer normal given the corresponding depth map. The architecture of AlGaN involves a two-stage CNN. The first stage infers the normal map using point cloud as input. If the light source position and grayscale image can be provided further, the second stage infers the albedo altogether with the normals from the first stage. In addition, the loss function is based on least-square error and Lambertian reflection.

With the help of synthetic data in Unity, a dataset is created for CNN model training. It can provide accurate ground truth for training work, which real data is usually not provided. The results of this dataset show that our AlGaN model performs also well on real data captured by Infrared cameras. The trained Normal models achieve a remarkably better prediction accuracy at a low computational cost compared to the standard approaches for semi-dense point clouds.

>>>>

Surface normal is an important property of a surface with many applications, like surface reconstruction, shadings generation and other visual effects. The quality of the surface normal is directly impact the performances of the techniques mentioned above. However, especially in the task of real-world object digitalization, the surface is usually hardly to be mathematically described in equations due to the elaborate details on the objects. Instead, it is common to use a group of points to describe the object surface, which is a memory economical solution and also can be easier measured by 3D scanners.  The working principle of scanners are various due to the application scenarios, which consequently produce different point cloud structure. For the scanners without positions recording, the point cloud acquired after scanning is unstructured. In this case, every 3D point can be captured by different capture position, and neighbors are not defined by capture time, which increase the difficulty and computation for the neighbor based normal inference approaches.  Furthermore, since the lack of inherent structure, the normal can hardly be inferred by the parallel approaches. 
To acquire the structured point cloud, a depth camera can be used for data collection. It captures the RGB-D images for the object, which includes the standard RGB image with depth information of each pixel. After camera calibration, the corresponding point cloud can be calculated based on the depth map and camera matrix. The advantages are, a structured point cloud is mapped directly based on the same capture position from a 2D depth map, the neighbor information of each point is identical to the corresponding pixel in the 2D depth map, which reduced a huge computational works. Besides, base on the RGB image, which has the same indices with depth map, the normal inference task can also utilize the relationship between surface normal and light reflection to further rectify the inferred normal map with knowing light source. Furthermore, the structured point cloud has the same shape with the standard RGB images, which can be used for deep learning based approaches. 
Unfortunately, the depth maps captured by the sensors are only semi-dense, which is mainly caused by optical noises and the reflections in dark and shinny areas. Consequently, it raises the challenges to the accuracy of normal inference tasks. 
Recently, deep learning-based methods are used in computer vision tasks such as image segmentation, image inpainting, and depth density enhancement. The output of the model is usually has the same shape as the input but with different content, which is by comparison similar to the normal inference task. 
The deep learning based method has many advantages, first, it deals well with semi-dense input, which is a common issue for the depth map. Second, the network of deep learning model handles the structured point cloud as a single input and infer the corresponding normal all together, which is very time economical comparing to the approaches like neighbor based methods. Furthermore, if the texture image acquiring from the camera is already illuminated by strong directional light of a video projector, whose position is known, then there should exist theoretical relations among light direction, normal direction, and grayscale image. Then the normal can be inferred better using the given image information and depth map.  Based on the grayscale image and corresponding semi-dense depth maps, a multi-staged model can be designed to infer the normal map, which can give more density and robust results comparing to the standard algorithm.
In this work, we apply the deep learning based approach for surface normal inference, which is based on the calibrated illuminated RGB-D image. A network named gated convolution neural network (GCNN) is proposed for normal inference, which is robust to the semi-dense input. 
With the help of synthetic data in Unity, a dataset is also created for CNN model training. It can provide accurate ground truth for training work, which real data is usually not provided. The trained model is performed well both in synthetic dataset as well as the real dataset captured from RGB-D cameras and achieve a remarkably better prediction accuracy at a low computational cost compared to the standard approaches for semi-dense point clouds.  
The structure of the thesis is as follows, Chapter 1 is the introduction of the whole work. Chapter 2 briefly discusses the related work about normal inference. Chapter 3 is the main approaches of this work. Chapter 4 introduces the created dataset for the training work. Chapter 5 is the description of the experiments and the evaluation of the models. Chapter 6 is the conclusion of the whole thesis.

### 2022-06-26 paper/Chapters/Chapter02.tex
update ag model

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

>>>>

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\section{Standard Methods}

%%%%%%%%%%%%%%%%%%% talk about the reason of this thesis %%%%%%%%%%%%%%%%%%

%% standard method for normal inference is insufficient, input is sparse, not robust enough
Standard methods compute normals from the point cloud using neighboring information in image space or from a single grayscale image using use Shape from Shading \cite{SFS}. The first method assumes that the neighbors of the points locate on the same plane. This method performs well with a well-chosen window size. However, the drawbacks are that the algorithm is highly noise sensitive. It is weak in handling missing pixels, which is a common issue in the input data. The second method depends on the correct information about the light source. Errors may occur in regions with inter-reflections in the 3D measurement.

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\section{Deep Learning based Method}
%% CNN based methods
Recently, deep learning based method \cite{yolov3}, \cite{efficientDet} achieved a great succeed for image processing. 
These network architectures use a batch of RGB/Grayscale images as input and usually employ for classification problems. Usually, the images are convoluted with a convolutional layer and downsampling with pooling layers. The outputs of the network consist of a single value to represent the index of the corresponding class \cite{efficientDet}, or with a set of values to represent the position of bounding boxes.\cite{yolov3}.

However, in many other vision tasks, like normal map inference, the output is demanded as the same shape as the input. Instead of predicting one or several classes for the whole input matrix, the class for each pixel requires for prediction. In this case, the traditional network architecture is not suitable anymore.

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------
